---
title: "Using the ADF&G Flextable Helpers"
output: word_document
---

This vignette demonstrates the shared table helpers with fictional data. It does
not read report inputs or require access to the subsistence data.

The repository currently uses these functions by sourcing an R file; it is not
an installable R package. The example is reproducible from a repository
checkout, provided the required R packages are installed.

```{r setup, include=FALSE}
required_packages <- c("flextable", "officer", "knitr")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages) > 0) {
  stop("Install required packages first: ", paste(missing_packages, collapse = ", "))
}

helper_candidates <- c(
  file.path("functions", "adfg_flextable_helpers.R"),
  file.path("..", "functions", "adfg_flextable_helpers.R")
)
existing_helpers <- helper_candidates[file.exists(helper_candidates)]
if (length(existing_helpers) == 0L) {
  stop("Run this vignette from the repository root or its best_practices directory.")
}
source(normalizePath(existing_helpers[[1]], winslash = "/"), local = knitr::knit_global())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Create a table

Start with an ordinary data frame. The labels and estimates below are
illustrative only.

```{r build-table}
example_data <- data.frame(
  Community = c("Koyuk", "Nulato", "Total"),
  Harvest = c(1250, 980, 2230),
  stringsAsFactors = FALSE
)
numeric_cols <- "Harvest"

ft <- flextable::flextable(example_data)
ft <- adfg_ft_apply_table_guardrails(
  ft,
  left_cols = "Community",
  right_cols = numeric_cols,
  header_center_cols = numeric_cols,
  font_size = 9,
  word_title = "Illustrative subsistence harvest table",
  word_description = "Illustrative harvest estimates for three rows."
)
ft <- adfg_ft_apply_context_borders(
  ft,
  row_labels = example_data$Community,
  numeric_cols = numeric_cols
)
```

`adfg_ft_apply_table_guardrails()` applies the shared typography, alignment,
header rules, width limits, and Word accessibility properties.
`adfg_ft_apply_context_borders()` adds a separator for the total row.

## Add notes and render

Footer notes use an explicit note-to-marker mapping. Printing the resulting
flextable includes it in this Word document.

```{r footer-notes}
note <- "All values in this example are fictional."
ft <- adfg_ft_add_footer_notes(
  ft,
  note_lines = note,
  marker_map = stats::setNames(list("a"), note),
  base_note = "ND indicates no data."
)
ft
```

## Run it

From the repository root, render this document with:

```r
rmarkdown::render(
  "best_practices/adfg_flextable_helpers_vignette.Rmd",
  output_file = "adfg_flextable_helpers_vignette.docx",
  output_dir = tempdir()
)
```

The helper source is [functions/adfg_flextable_helpers.R](../functions/adfg_flextable_helpers.R).
The standalone script in [examples/adfg_flextable_helpers_example.R](../examples/adfg_flextable_helpers_example.R)
creates the same kind of DOCX directly with `Rscript`.